\section*{Abstract}
\addcontentsline{toc}{chapter}{Abstract}
In recent years, the use of cloud technologies is rapidly becoming more common as people realize the highly scalable, configurable, efficient and on-demand remote access to computing resources which is available to them via the cloud. People use cloud services to accomplish a variety of important tasks, like developing and accessing cloud-native applications, carrying out large-scale data analytics or conducting high-performance computing and much more. One particularly popular use case involves the use of cloud technologies for storage, thereby allowing people to, for instance, satisfy the storage needs of their applications, archive and backup important data, as well as synchronize and share files with others to collaborate in real time. In our day and age, there exist many different cloud storage services which differ in many aspects like performance, security or capabilities. However, because each service has its own specific interface, an issue arises when users want to adopt more than one service at a time. In cases like these, users are required to learn and regularly switch between different interfaces, which can easily become very cumbersome. The purpose of this project, therefore, is to provide a common interface for accessing many different cloud storage services. We accomplish this by developing a Python library that makes use of and integrates the official SDKs of four popular such services (Dropbox, Box, Google Drive and Amazon S3) for upload, download and deletion of files, while laying the groundwork for the implementation of additional services and functionalities. We also measure the execution time of our library when working on various workloads to evaluate the overall performance of our library.